---
title: "statistics experience"
author: "Helene Gu"
date: "03/06/2022"
output:
  pdf_document: default
  html_document: default
---

## load libraries and data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r packages, message = F, warning = F}
library(tidyverse)
library(viridis)
```
```{r load-data, message = FALSE, warning = F, show_col_types = FALSE }
stations <- read_csv("~/Labs/tidytuesday/data/2022/2022-03-01/stations.csv")
```



## Tidy Tuesday challenge 03/01

```{r remove-planned-and-temporarily-unavailable}
#separate planned/temporarily unavailable
stations2 <- stations%>%
   filter(!str_detect(GROUPS_WITH_ACCESS_CODE, "TEMPORARILY UNAVAILABLE"))%>%
   filter(!str_detect(GROUPS_WITH_ACCESS_CODE, "PLANNED"))

```
```{r planned}
#planned
stations3 <- stations%>%
   filter(str_detect(GROUPS_WITH_ACCESS_CODE, "PLANNED"))%>%
  mutate(PLANNEDYES = "planned")

```

```{r visualization}
#visualization
stations2%>%
  group_by(CITY)%>%
  ggplot() +
  geom_point(data = stations2, aes(x = LONGDD, y = LATDD,color=STATE),
             size=0.0001) +
  geom_point(data = stations3, aes(x = LONGDD, y = LATDD),color="black",
             size=0.0001)+
  labs(title = "Are we putting alternative energy stations where we lack them?",
       subtitle = "Where black dots are planned energy stations",
         x= "Longitude",
       y="Latitude")+
  guides(color = "none")

```
States that lack alternative energy stations include Arkansas, Montana, Wyoming,Nevada, and Delaware. This can be seen by the sparser dots in the visualization.Arkansas and Delaware will not be getting any new stations, Montana and Wyoming will be getting 2,and Nevada 3.This is compared to 73 new stations in CA, 28 in FL and 22 in NY. It doesn't look like the planned stations are going where there is a geographic lack of stations, but instead 
mostly to states where there already are a lot of them. 

## Analysis
```{r analysis}
#states with least alternative energy stations
stations2%>%
  group_by(STATE)%>%
  count(n())%>%
  arrange((n))

#states with most planned alternative energy stations
stations3%>%
  group_by(STATE)%>%
  count(n())%>%
  arrange(desc(n))
```
credits : 

https://stackoverflow.com/questions/34444295/how-to-specify-does-not-contain-in-dplyr-filter
https://www.earthdatascience.org/courses/earth-analytics/spatial-data-r/make-maps-with-ggplot-in-R/
